---
title: "ALV_PNAS_Reproduction"
output: html_document
date: "2026-02-12"
---

##SET-UP 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# -------------------------------
# 0) Package loader
# -------------------------------
load_pkg <- function(pkg) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg, dependencies = TRUE)
  }
  suppressPackageStartupMessages(library(pkg, character.only = TRUE))
}

pkgs <- c(
  "here","haven","dplyr","tidyr","stringr","stringi",
  "sandwich","lmtest","MASS","broom","ordinal",
  "ggplot2","gridExtra","grid","tools","purrr","tibble"
)
invisible(lapply(pkgs, load_pkg))

# -------------------------------
# 1) Define required file locations
# -------------------------------
dir_data_raw <- here::here("data", "raw")

required_files <- c("OriginalData.dta", "NewAdidaCoding.dta", "SurveyData.dta")
required_paths <- file.path(dir_data_raw, required_files)

missing <- required_files[!file.exists(required_paths)]
if (length(missing) > 0) {
  stop(
    "Missing required data file(s) in data/raw/: ",
    paste(missing, collapse = ", "),
    "\n\nFix: Put these files in: ", dir_data_raw,
    "\nThen re-run the Rmd."
  )
}

# -------------------------------
# 2) Output directories
# -------------------------------
dir_results <- here::here("results")
dir_tables  <- file.path(dir_results, "tables")
dir_figures <- file.path(dir_results, "figures")

dir.create(file.path(dir_tables, "cv"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(dir_tables, "sv"), recursive = TRUE, showWarnings = FALSE)
dir.create(dir_figures, recursive = TRUE, showWarnings = FALSE)

# -------------------------------
# 3) Read Stata .dta files (with Latin-1 encoding for French characters)
# -------------------------------
original_data <- haven::read_dta(file.path(dir_data_raw, "OriginalData.dta"), encoding = "latin1")
newcoding     <- haven::read_dta(file.path(dir_data_raw, "NewAdidaCoding.dta"), encoding = "latin1")
survey_data   <- haven::read_dta(file.path(dir_data_raw, "SurveyData.dta"), encoding = "latin1")

# -------------------------------
# 4) Output collectors
# -------------------------------
cv_out <- list()
sv_out <- list()

```

##Creating output pathaways: 

```{r}

# Save a data frame as one or more PNG pages
save_table_image_pages <- function(df, filename_stub, title,
                                   out_dir = "tables",
                                   rows_per_page = 30,
                                   width = 2000, height = 1400, res = 200,
                                   digits = 4) {
  dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

  # Coerce to data.frame 
  df2 <- as.data.frame(df)

  # Round numeric columns 
  df2 <- df2 %>%
    mutate(across(where(is.numeric), ~ round(.x, digits)))

  # Split into pages
  n <- nrow(df2)
  n_pages <- ceiling(n / rows_per_page)

  out_files <- character(0)

  for (p in seq_len(n_pages)) {
    idx_start <- (p - 1) * rows_per_page + 1
    idx_end   <- min(p * rows_per_page, n)
    df_page   <- df2[idx_start:idx_end, , drop = FALSE]

    fn <- file.path(out_dir, paste0(filename_stub, "_p", p, ".png"))
    out_files <- c(out_files, fn)

    png(fn, width = width, height = height, res = res)
    grid::grid.newpage()

    page_title <- if (n_pages > 1) {
      paste0(title, " (page ", p, " of ", n_pages, ")")
    } else title

    grid::grid.text(page_title, x = 0.02, y = 0.97, just = "left",
                    gp = grid::gpar(fontsize = 16, fontface = "bold"))

    # Table
    gridExtra::grid.table(df_page, rows = NULL)
    dev.off()
  }

  invisible(out_files)
}

##Helper: save everything inside a list (cv_out / sv_out)

save_all_tables_from_list <- function(out_list, prefix,
                                      base_dir = "tables",
                                      rows_per_page = 30,
                                      width = 2000, height = 1400, res = 200,
                                      digits = 4) {

  out_dir <- file.path(base_dir, prefix)
  dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

  # Save an index 
  index <- tibble::tibble(
    object_name = names(out_list),
    n_rows = purrr::map_int(out_list, ~ nrow(as.data.frame(.x))),
    n_cols = purrr::map_int(out_list, ~ ncol(as.data.frame(.x)))
  )

  # Write index image too
  save_table_image_pages(
    df = index,
    filename_stub = paste0(prefix, "_INDEX"),
    title = paste0(toupper(prefix), " output index"),
    out_dir = out_dir,
    rows_per_page = rows_per_page,
    width = width, height = height, res = res,
    digits = 0
  )

  # Save each element
  for (nm in names(out_list)) {
    df <- out_list[[nm]]

    # Make filename
    safe_nm <- gsub("[^A-Za-z0-9_\\-]+", "_", nm)

    save_table_image_pages(
      df = df,
      filename_stub = safe_nm,
      title = paste0(prefix, ": ", nm),
      out_dir = out_dir,
      rows_per_page = rows_per_page,
      width = width, height = height, res = res,
      digits = digits
    )
  }

  invisible(TRUE)
}

```


##PART 1 — CV (RÉSUMÉ AUDIT) EXPERIMENT
#  Original Stata file: 2009Dec17CLA_CV.do
#   Design: matched pairs of CVs sent to employers.
#    • "Test" candidate  = candidate 1 (Khadija or Marie)
#    • "Control" candidate = candidate 2 (Aurélie)
#    • MuslimPair = 1 when pair is Khadija vs Aurélie
#    • ChristianPair = 1 when pair is Marie vs Aurélie

#  Three codings of the outcome differ in what counts as "missing":
# Coding 1 (ISM-Corum): drop NRNR, NRNo, NoNR, NoNo, ""
# Coding 2: drop NRNR, NRNo, NoNR, "" (keep NoNo as 0)
# Coding 3: drop only "" (invalid email tests)

# ── 1a. Variable construction ────────────────────────────────────────────────
# Stata generates binary indicators from string variables for pair type, position, contract, and firm size.  
# To replicate this in R: case_when

```{r}


cv <- original_data %>%
  mutate(
    # --- normalize firm size ---
    firm_size = firm_size %>%
      as.character() %>%
      str_replace_all("\u00A0", " ") %>%
      str_squish() %>%
      stri_replace_all_regex("\\p{Cc}", "") %>%   
      stri_trans_general("Latin-ASCII") %>%
      na_if(""),

    # --- normalize position ---
    position_clean = position %>%
      as.character() %>%
      str_replace_all("\u00A0", " ") %>%
      str_squish() %>%
      stri_replace_all_regex("\\p{Cc}", "") %>%   
      stri_trans_general("Latin-ASCII") %>%
      na_if(""),

    # --- pull the position code (01/02/03/04) ---
    position_code = str_extract(position_clean, "^\\d{2}"),

    # Pair type
    MuslimPair    = as.integer(str_detect(pair, "Khadija")),
    ChristianPair = as.integer(str_detect(pair, "Marie")),

    # Job position dummies 
    Secretary           = as.integer(position_code == "01"),
    AssistantDirector   = as.integer(position_code == "02"),
    AssistantAccountant = as.integer(position_code == "03"),
    Accountant          = as.integer(position_code == "04"),

    # Contract type
    CDD = as.integer(contract_type == "CDD"),
    CDI = as.integer(contract_type == "CDI"),

    # Firm size dummies
    SmallFirm  = case_when(
      is.na(firm_size) ~ NA_integer_,
      str_starts(firm_size, "01_") ~ 1L,
      TRUE ~ 0L
    ),
    MediumFirm = case_when(
      is.na(firm_size) ~ NA_integer_,
      str_starts(firm_size, "02_") ~ 1L,
      TRUE ~ 0L
    ),
    LargeFirm  = case_when(
      is.na(firm_size) ~ NA_integer_,
      str_starts(firm_size, "03_") ~ 1L,
      TRUE ~ 0L
    )
  )

```


# ── 1b. Merge corrected result variable ──────────────────────────────────────
# Stata: joinby testnumber using NewAdidaCoding.dta
# The corrected "newresult" variable fixes errors in the original "result".

```{r}

cv <- cv %>%
  mutate(testnumber = str_trim(as.character(testnumber)))

newcoding <- newcoding %>%
  mutate(testnumber = str_trim(as.character(testnumber)))

cvj <- left_join(cv, newcoding, by = "testnumber") %>%
  mutate(newresult = str_trim(as.character(newresult)))


```

#── 1c. Helpers: pseudo-R2, robust logit, & tidy output ──────────────────────────────────
# Stata: logit Y X, robust  →  uses Huber-White (HC0) standard errors.

```{r}

tidy_rlogit <- function(formula, data, model_name) {
  mf <- model.frame(formula, data = data, na.action = na.omit) # exact estimation sample
  fit <- glm(formula, data = mf, family = binomial(link = "logit"))
  ct  <- lmtest::coeftest(fit, vcov. = sandwich::vcovHC(fit, type = "HC0"))

  # Fit null model explicitly on the same sampled data
  null_fit <- glm(update(formula, . ~ 1), data = mf, family = binomial(link = "logit"))

  ll_full <- as.numeric(logLik(fit))
  ll_null <- as.numeric(logLik(null_fit))
  pseudo_r2 <- 1 - (ll_full / ll_null)

  tibble::tibble(
    model     = model_name,
    nobs      = nobs(fit),
    pseudo_r2 = pseudo_r2,
    term      = rownames(ct),
    estimate  = ct[, "Estimate"],
    std.error = ct[, "Std. Error"],
    statistic = ct[, "z value"],
    p.value   = ct[, "Pr(>|z|)"]
  )
}

```

# ── 1d. Helper:  ───────────────────────────
# Returns the data with ControlWins, TestWins, and the interact term.

```{r}

build_outcomes <- function(data, missing_values) {
  data %>%
    mutate(
      ControlWins = case_when(
        newresult %in% c("NRYes", "NoYes", "YesYes") ~ 1,
        newresult %in% missing_values                 ~ NA_real_,
        TRUE                                          ~ 0),
      TestWins = case_when(
        newresult %in% c("YesNR", "YesNo", "YesYes") ~ 1,
        newresult %in% missing_values                 ~ NA_real_,
        TRUE                                          ~ 0),
      interact = MuslimPair * test_photo
    )
}

```

# ── 1e. Define the three codings ────────────────────────────────────────────

```{r}

coding_defs <- list(
  C1 = c("", "NRNR", "NRNo", "NoNR", "NoNo"),   
  C2 = c("", "NRNR", "NRNo", "NoNR"),            
  C3 = c("")                                      
)

```

# ── 1f. Run all analyses across the three codings ───────────────────────────
# For each coding we produce:
#   Summary statistics (overall, by MuslimPair, by ChristianPair)
#   Welch t-tests (TestWins by MuslimPair; ControlWins by MuslimPair)
#   10 logit models with robust SEs 

```{r}

for (cname in names(coding_defs)) {

  d      <- build_outcomes(cvj, coding_defs[[cname]])
  
  # Drop rows with NA in test_photo before fitting any models!
  d <- d %>% filter(!is.na(test_photo))
  
  suffix <- cname   

  d      <- build_outcomes(cvj, coding_defs[[cname]])
  suffix <- cname   

  # Summary statistics
stats <- list(
  n    = ~ sum(!is.na(.)),
  mean = ~ mean(., na.rm = TRUE),
  sd   = ~ sd(., na.rm = TRUE)
)

overall <- d %>%
  summarise(
    across(c(TestWins, ControlWins), stats, .names = "{.col}_{.fn}")) %>%
  mutate(pair = "All", .before = 1)

by_pair <- bind_rows(
  d %>% filter(MuslimPair    == 1) %>%
    summarise(pair = "Muslim",    across(c(TestWins, ControlWins), stats, .names = "{.col}_{.fn}")),
  d %>% filter(ChristianPair == 1) %>%
    summarise(pair = "Christian", across(c(TestWins, ControlWins), stats, .names = "{.col}_{.fn}"))
)

cv_out[[paste0("summary_", suffix)]] <- bind_rows(overall, by_pair)

  # ---- t-tests ----
  # Stata: ttest TestWins, by(MuslimPair) unpaired unequal
  # Stata: ttest ControlWins, by(MuslimPair) unpaired unequal
  # Stata: ttest TestWins==ControlWins if ChristianPair==1, unpaired unequal
  # Stata: ttest TestWins==ControlWins if MuslimPair==1, unpaired unequal

  d_chr <- d %>% filter(ChristianPair == 1, !is.na(TestWins), !is.na(ControlWins))
  d_mus <- d %>% filter(MuslimPair    == 1, !is.na(TestWins), !is.na(ControlWins))

  tt <- bind_rows(
    tidy(t.test(TestWins    ~ MuslimPair, data = d, var.equal = FALSE)) %>%
      mutate(test = "TestWins by MuslimPair"),
    tidy(t.test(ControlWins ~ MuslimPair, data = d, var.equal = FALSE)) %>%
      mutate(test = "ControlWins by MuslimPair"),
    tidy(t.test(d_chr$TestWins, d_chr$ControlWins, paired = FALSE, var.equal = FALSE)) %>%
      mutate(test = "Test==Control | Christian"),
    tidy(t.test(d_mus$TestWins, d_mus$ControlWins, paired = FALSE, var.equal = FALSE)) %>%
      mutate(test = "Test==Control | Muslim")
  )

  cv_out[[paste0("ttests_", suffix)]] <- tt

# ---- Logit models ----

tidy_rlogit_m9_stata <- function(data, suffix) {

  data <- data %>% mutate(interact = MuslimPair * test_photo)

  f_full <- ControlWins ~ MuslimPair + CDD + SmallFirm + MediumFirm +
    Secretary + AssistantDirector + AssistantAccountant + test_photo + interact

  d_int <- data %>% filter(!is.na(interact), interact != 0, !is.na(ControlWins))
  sep_hit <- nrow(d_int) > 0 && (all(d_int$ControlWins == 1) || all(d_int$ControlWins == 0))

  # helper: McFadden pseudo-R2 on same estimation sample as fit
  # null_fit should use exact same model.frame as fit
  mcfadden_same_sample <- function(fit) {
    # Extract the exact estimation sample used by fit (accounts for all NA dropping)
    mf <- model.frame(fit)
    # Fit null model on the same sample, not on original data
    null_fit <- glm(update(formula(fit), . ~ 1), 
                    data = mf, 
                    family = binomial(link = "logit"))
    
    ll_full  <- as.numeric(stats::logLik(fit))
    ll_null  <- as.numeric(stats::logLik(null_fit))
    1 - (ll_full / ll_null)
  }

  if (sep_hit) {
    # drop interact term & drop interact!=0 observations
    d_fit <- data %>% filter(is.na(interact) | interact == 0)

    f_drop <- ControlWins ~ MuslimPair + CDD + SmallFirm + MediumFirm +
      Secretary + AssistantDirector + AssistantAccountant + test_photo

    # Create model.frame first with all variables in formula
    # Just to ensure both full and null models see identical sample
    mf_fit <- model.frame(f_drop, data = d_fit, na.action = na.omit)
    
    fit <- suppressWarnings(glm(f_drop, 
                                data = mf_fit, 
                                family = binomial(link = "logit")))
    ct  <- lmtest::coeftest(fit, vcov. = sandwich::vcovHC(fit, type = "HC0"))

    # McFadden on same sample
    pseudo_r2 <- mcfadden_same_sample(fit)

    res <- tibble::tibble(
      model     = paste0(suffix, "_m9_CW_interact"),
      nobs      = stats::nobs(fit),
      pseudo_r2 = pseudo_r2,
      term      = rownames(ct),
      estimate  = ct[, "Estimate"],
      std.error = ct[, "Std. Error"],
      statistic = ct[, "z value"],
      p.value   = ct[, "Pr(>|z|)"],
      note      = NA_character_
    ) %>%
      dplyr::bind_rows(
        tibble::tibble(
          model     = paste0(suffix, "_m9_CW_interact"),
          nobs      = stats::nobs(fit),
          pseudo_r2 = pseudo_r2,
          term      = "interact",
          estimate  = NA_real_, std.error = NA_real_,
          statistic = NA_real_, p.value = NA_real_,
          note      = "dropped: interact!=0 obs removed"
        )
      )

    return(res)
  }

  # full model branch
  
  mf_fit <- model.frame(f_full, data = data, na.action = na.omit)
  
  fit <- suppressWarnings(glm(f_full, 
                              data = mf_fit, 
                              family = binomial(link = "logit")))
  ct  <- lmtest::coeftest(fit, vcov. = sandwich::vcovHC(fit, type = "HC0"))

  pseudo_r2 <- mcfadden_same_sample(fit)

  tibble::tibble(
    model     = paste0(suffix, "_m9_CW_interact"),
    nobs      = stats::nobs(fit),
    pseudo_r2 = pseudo_r2,
    term      = rownames(ct),
    estimate  = ct[, "Estimate"],
    std.error = ct[, "Std. Error"],
    statistic = ct[, "z value"],
    p.value   = ct[, "Pr(>|z|)"],
    note      = NA_character_
  )
}

tidy_rlogit <- function(formula, data, model_name) {
  mf <- model.frame(formula, data = data, na.action = na.omit)
  # Fit full model on this controlled sample
  fit <- glm(formula, data = mf, family = binomial(link = "logit"))
  ct  <- lmtest::coeftest(fit, vcov. = sandwich::vcovHC(fit, type = "HC0"))
  # Fit null model on the same sample (not on original data!)
  null_fit <- glm(update(formula, . ~ 1), 
                  data = mf, 
                  family = binomial(link = "logit"))

  # McFadden pseudo-R2: both LLs computed on same N
  ll_full <- as.numeric(stats::logLik(fit))
  ll_null <- as.numeric(stats::logLik(null_fit))
  pseudo_r2 <- 1 - (ll_full / ll_null)

  tibble::tibble(
    model     = model_name,
    nobs      = stats::nobs(fit),
    pseudo_r2 = pseudo_r2,
    term      = rownames(ct),
    estimate  = ct[, "Estimate"],
    std.error = ct[, "Std. Error"],
    statistic = ct[, "z value"],
    p.value   = ct[, "Pr(>|z|)"]
  )
}

logit_results <- dplyr::bind_rows(

  tidy_rlogit(ControlWins ~ MuslimPair + CDD + SmallFirm + MediumFirm,
              d, paste0(suffix, "_m1_CW_base")),
  tidy_rlogit(ControlWins ~ MuslimPair + CDD + SmallFirm + MediumFirm + test_photo,
              d, paste0(suffix, "_m2_CW_photo")),

  tidy_rlogit(TestWins ~ MuslimPair + CDD + SmallFirm + MediumFirm,
              d, paste0(suffix, "_m3_TW_base")),
  tidy_rlogit(TestWins ~ MuslimPair + CDD + SmallFirm + MediumFirm + test_photo,
              d, paste0(suffix, "_m4_TW_photo")),

  tidy_rlogit(ControlWins ~ MuslimPair + CDD + SmallFirm + MediumFirm +
                Secretary + AssistantDirector + AssistantAccountant,
              d, paste0(suffix, "_m5_CW_pos")),
  tidy_rlogit(ControlWins ~ MuslimPair + CDD + SmallFirm + MediumFirm +
                Secretary + AssistantDirector + AssistantAccountant + test_photo,
              d, paste0(suffix, "_m6_CW_pos_photo")),

  tidy_rlogit(TestWins ~ MuslimPair + CDD + SmallFirm + MediumFirm +
                Secretary + AssistantDirector + AssistantAccountant,
              d, paste0(suffix, "_m7_TW_pos")),
  tidy_rlogit(TestWins ~ MuslimPair + CDD + SmallFirm + MediumFirm +
                Secretary + AssistantDirector + AssistantAccountant + test_photo,
              d, paste0(suffix, "_m8_TW_pos_photo")),

  tidy_rlogit_m9_stata(d, suffix),

  tidy_rlogit(TestWins ~ MuslimPair + CDD + SmallFirm + MediumFirm +
                Secretary + AssistantDirector + AssistantAccountant +
                test_photo + interact,
              d, paste0(suffix, "_m10_TW_interact"))
)

cv_out[[paste0("logits_", suffix)]] <- logit_results

cv_out[[paste0("logit_fit_", suffix)]] <- logit_results %>%
  dplyr::distinct(model, nobs, pseudo_r2) %>%
  dplyr::arrange(model)}

  
```

#PART 2 — SURVEY RESULTS 
#  Original Stata file: SurveyDataForPNAS.do
#  Three analyses:
#    A. OLS regression of income on first-migrant education, self-reported
#       religion, gender, age  (robust SEs)
#    B. "Table 2" ordered probit (david_religion, with interaction)
#    C. "Table A4" ordered probit + predicted probabilities for
#       typical Muslim vs typical Christian

# ── 2a. Variable construction ───────────────────────


```{r}

sv <- survey_data %>%
  mutate(
    # Gender: 0 = male, 1 = female (Stata recodes sexe 1→0, 2→1)
    gender = case_when(sexe == 1 ~ 0L, sexe == 2 ~ 1L, TRUE ~ NA_integer_),
    female = as.integer(gender == 1),
    Age    = as.numeric(age1),

    # Education of first migrant (99 = missing in Stata)
    educ_firstmigrant = if_else(q15 == 99, NA_real_, as.numeric(q15)),

    # Self-reported religion: 0 = Muslim, 1 = Christian
    self_religion     = case_when(q53 == 1 ~ 0, q53 == 2 ~ 1, TRUE ~ NA_real_),

    # Income (99 = missing)
    income     = if_else(q95 == 99, NA_real_, as.numeric(q95)),
    income_ord = ordered(if_else(q95 == 99, NA_integer_, as.integer(q95))),

    # "david_religion": both family & self must agree
    david_religion = case_when(
      q7 == 1 & q53 == 1 ~ 0,
      q7 == 2 & q53 == 2 ~ 1,
      TRUE ~ NA_real_),

    # Family religion
    christianfamily = case_when(q7 == 2 ~ 1, q7 == 1 ~ 0, TRUE ~ NA_real_),

    # Head-of-household gender
    genderhead = case_when(
      (q40 == "01" & female == 0) | q40 == "03" ~ 1,
      (q40 == "01" & female == 1) | q40 == "04" ~ 0,
      TRUE ~ NA_real_),

    # Education variables
    curreduc   = if_else(q43 < 99, as.numeric(q43), NA_real_),
    educfather = if_else(q36 == 99, NA_real_, as.numeric(q36)),
    educmother = if_else(q37 == 99, NA_real_, as.numeric(q37)),

    # Education of household head
    educhead = case_when(
      q40 == "01" ~ curreduc,
      q40 == "03" ~ educfather,
      q40 == "04" ~ educmother,
      TRUE ~ NA_real_),

    initialeduc = if_else(q15 < 99, as.numeric(q15), NA_real_)
  )

#helper functions: 

pseudo_r2_mcfadden_clm <- function(fit) {
  # use same estimation sample
  mf <- model.frame(fit)
  y  <- all.vars(formula(fit))[1]
  null_form <- stats::as.formula(paste0(y, " ~ 1"))
  null_fit  <- ordinal::clm(null_form, data = mf, link = "probit")

  1 - as.numeric(logLik(fit)) / as.numeric(logLik(null_fit))
}

# Robust SEs for ordinal::clm 
vcov_robust_clm_stata <- function(fit) {
  # Score contributions (n x k)
  S <- as.matrix(sandwich::estfun(fit))

  # Hessian of the log-likelihood (k x k)
  H <- fit$Hessian

  B <- solve(H)
  meat <- crossprod(S)
  V <- B %*% meat %*% B
  V
}

tidy_robust_clm_stata <- function(fit, model_name) {
  V  <- vcov_robust_clm_stata(fit)
  ct <- lmtest::coeftest(fit, vcov. = V)

  tibble::tibble(
    model     = model_name,
    term      = rownames(ct),
    estimate  = ct[, 1],
    std.error = ct[, 2],
    statistic = ct[, 3],
    p.value   = ct[, 4]
  )
}

```

# ── 2b. OUTPUT A: OLS regression ────────────────────────────────────────────
# Stata: reg income educ_firstmigrant self_religion gender Age, robust

```{r}

m_ols      <- lm(income ~ educ_firstmigrant + self_religion + gender + Age, data = sv)
ols_robust <- tidy(coeftest(m_ols, vcov. = vcovHC(m_ols, type = "HC0")))
sv_out[["OLS_robust"]] <- ols_robust

sv_out[["OLS_summary_stats"]] <- sv %>%
  summarise(across(c(income, educ_firstmigrant, self_religion, gender, Age),
                   list(n    = ~ sum(!is.na(.)),
                        mean = ~ mean(., na.rm = TRUE),
                        sd   = ~ sd(., na.rm = TRUE)),
                   .names = "{.col}_{.fn}"))

ols_fit_stats <- tibble(
  model      = "OLS_robust",
  n_obs      = nobs(m_ols),
  r2         = summary(m_ols)$r.squared,
  adj_r2     = summary(m_ols)$adj.r.squared
)

sv_out[["OLS_robust"]] <- tidy(coeftest(m_ols, vcov. = vcovHC(m_ols, type = "HC0"))) %>%
  mutate(n_obs = nobs(m_ols))

```

# ── 2c. OUTPUT B: Table 2 — Ordered probit (david_religion) ─────────────────
# Stata:
#   oprobit q95 q15 gender age1 david_religion if q95~=99 & q15~=99, robust
#   gen interaction = david_religion * q15
#   oprobit q95 q15 gender age1 david_religion interaction if …, robust

```{r}

sv_t2 <- sv %>%
  dplyr::filter(!is.na(income_ord), !is.na(educ_firstmigrant)) %>%
  dplyr::mutate(interaction = david_religion * educ_firstmigrant)

op1 <- ordinal::clm(
  income_ord ~ educ_firstmigrant + gender + Age + david_religion,
  data = sv_t2, link = "probit"
)

op2 <- ordinal::clm(
  income_ord ~ educ_firstmigrant + gender + Age + david_religion + interaction,
  data = sv_t2, link = "probit"
)

# robust coefficient tables
sv_out[["Table2_oprobit_robust"]] <- dplyr::bind_rows(
  tidy_robust_clm(op1, "Table2_col1"),
  tidy_robust_clm(op2, "Table2_col2_interaction")
)

# fit stats (N + McFadden pseudo-R2)
sv_out[["Table2_fit_stats"]] <- dplyr::bind_rows(
  tibble::tibble(
    model = "Table2_col1",
    n_obs = stats::nobs(op1),
    pseudo_r2_mcfadden = pseudo_r2_mcfadden_clm(op1)
  ),
  tibble::tibble(
    model = "Table2_col2_interaction",
    n_obs = stats::nobs(op2),
    pseudo_r2_mcfadden = pseudo_r2_mcfadden_clm(op2)
  )
)


```

# ── 2d. OUTPUT C: Table A4 — Ordered probit + predicted probabilities ────────
# Stata restricts to head-of-household respondents (q40 in {01, 03, 04}) and estimates: oprobit income christianfamily genderhead educhead initialeduc, robust
# Then computes predicted probabilities for:
#   (i)   typical Muslim vs. Christian by group-specific modes
#   (ii)  typical person (overall modes), varying only religion
#   (iii) typical person within ±0.5 SD of each covariate mean

```{r}

svA4 <- sv %>%
  filter(q40 %in% c("01", "03", "04")) %>%
  filter(!is.na(income), !is.na(christianfamily),
         !is.na(genderhead), !is.na(educhead), !is.na(initialeduc)) %>%
  mutate(income_ord = ordered(income))

opA4 <- polr(income_ord ~ christianfamily + genderhead + educhead + initialeduc,
             data = svA4, method = "probit", Hess = TRUE)

sv_out[["TableA4_oprobit"]] <- broom::tidy(opA4)
sv_out[["TableA4_fit_stats"]] <- tibble::tibble(
  model = "TableA4",
  n_obs = nobs(opA4),
  pseudo_r2_mcfadden = pseudo_r2_mcfadden(opA4)
)


# Predicted probabilities
pmat  <- predict(opA4, newdata = svA4, type = "probs")
pcols <- paste0("p", seq_len(ncol(pmat)))
svA4p <- bind_cols(svA4, as.data.frame(pmat) %>% setNames(pcols))

# Helper: statistical mode
Mode <- function(x) {
  x  <- x[!is.na(x)]
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# Helper: mean predicted probs for a subset
mean_probs <- function(df, label) {
  df %>%
    summarise(across(all_of(pcols), ~ mean(., na.rm = TRUE)), n = n()) %>%
    mutate(group = label, .before = 1)
}

# (i) Group-specific modes
modes_mus <- svA4p %>% filter(christianfamily == 0) %>%
  summarise(across(c(genderhead, educhead, initialeduc), Mode))
modes_chr <- svA4p %>% filter(christianfamily == 1) %>%
  summarise(across(c(genderhead, educhead, initialeduc), Mode))

pp_groupmode <- bind_rows(
  svA4p %>%
    filter(christianfamily == 0,
           genderhead  == modes_mus$genderhead,
           educhead    == modes_mus$educhead,
           initialeduc == modes_mus$initialeduc) %>%
    mean_probs("Muslim (group mode)"),
  svA4p %>%
    filter(christianfamily == 1,
           genderhead  == modes_chr$genderhead,
           educhead    == modes_chr$educhead,
           initialeduc == modes_chr$initialeduc) %>%
    mean_probs("Christian (group mode)")
)

# (ii) Overall modes, varying only religion
modes_all <- svA4p %>%
  summarise(across(c(genderhead, educhead, initialeduc), Mode))

pp_overallmode <- bind_rows(
  svA4p %>%
    filter(christianfamily == 0,
           genderhead  == modes_all$genderhead,
           educhead    == modes_all$educhead,
           initialeduc == modes_all$initialeduc) %>%
    mean_probs("Muslim (overall mode)"),
  svA4p %>%
    filter(christianfamily == 1,
           genderhead  == modes_all$genderhead,
           educhead    == modes_all$educhead,
           initialeduc == modes_all$initialeduc) %>%
    mean_probs("Christian (overall mode)")
)

# (iii) Mean ± 0.5 SD bands (overall)
band_filter <- function(df, ref_df = df) {
  stats <- ref_df %>%
    summarise(
      educhead_m    = mean(educhead, na.rm = TRUE),
      educhead_s    = sd(educhead, na.rm = TRUE),
      initialeduc_m = mean(initialeduc, na.rm = TRUE),
      initialeduc_s = sd(initialeduc, na.rm = TRUE)
    )

  df %>%
    filter(
      between(educhead,
              stats$educhead_m    - 0.5 * stats$educhead_s,
              stats$educhead_m    + 0.5 * stats$educhead_s),
      between(initialeduc,
              stats$initialeduc_m - 0.5 * stats$initialeduc_s,
              stats$initialeduc_m + 0.5 * stats$initialeduc_s)
    )
}

pp_band <- bind_rows(
  svA4p %>% filter(christianfamily == 0) %>% band_filter() %>% mean_probs("Muslim (±0.5 SD, group)"),
  svA4p %>% filter(christianfamily == 1) %>% band_filter() %>% mean_probs("Christian (±0.5 SD, group)")
)

pp_band_overall <- band_filter(svA4p, svA4p) %>%
  dplyr::group_by(christianfamily) %>%
  dplyr::summarise(dplyr::across(all_of(pcols), ~ mean(.x, na.rm = TRUE)),
                   n = dplyr::n(),
                   .groups = "drop") %>%
  dplyr::mutate(group = dplyr::if_else(christianfamily == 1,
                                      "Christian (±0.5 SD, overall)",
                                      "Muslim (±0.5 SD, overall)")) %>%
  dplyr::select(group, dplyr::everything(), -christianfamily)

sv_out[["TableA4_predprob_groupmode"]]    <- pp_groupmode
sv_out[["TableA4_predprob_overallmode"]]  <- pp_overallmode
sv_out[["TableA4_predprob_band_group"]]   <- pp_band
sv_out[["TableA4_predprob_band_overall"]] <- pp_band_overall

a4_fit_stats <- tibble(
  model = "TableA4",
  n_obs = nobs(opA4),
  logLik = as.numeric(logLik(opA4)),
  pseudo_r2_mcfadden = pseudo_r2_mcfadden(opA4)
)

svA4 <- sv %>%
  dplyr::filter(q40 %in% c("01", "03", "04")) %>%
  dplyr::filter(!is.na(income), !is.na(christianfamily),
                !is.na(genderhead), !is.na(educhead), !is.na(initialeduc)) %>%
  dplyr::mutate(income_ord = ordered(income))

opA4 <- ordinal::clm(
  income_ord ~ christianfamily + genderhead + educhead + initialeduc,
  data = svA4, link = "probit"
)

sv_out[["TableA4_oprobit_robust"]] <- tidy_robust_clm_stata(opA4, "TableA4")

```

# ── SAVING CV & SURVEY RESULTS  ──────────────────────

```{r}

save_all_tables_from_list(
  out_list = cv_out,
  prefix = "cv",
  base_dir = dir_tables,   
  rows_per_page = 20,
  width = 2400,
  height = 1600,
  res = 200,
  digits = 4
)

save_all_tables_from_list(
  out_list = sv_out,
  prefix = "sv",
  base_dir = dir_tables,   
  rows_per_page = 20,
  width = 2400,
  height = 1600,
  res = 200,
  digits = 4
)

cv_out <- cv_out[sort(names(cv_out))]
sv_out <- sv_out[sort(names(sv_out))]

```

# ── PART 3: FIGURE REPLICATION  ──────────────────────

```{r}

## FIGURE 1  (two-panel coef plot)
jpeg(
  filename = file.path(dir_figures, "PNAS2010ALVTable1.jpg"),
  width = 3700, height = 2000, quality = 100, res = 300
)

layout(matrix(c(1,2),1,2), widths=c(1,1))

margins  <- c(4, 6, 1, 0)  # bottom, left, top, right
margins2 <- c(4, 3, 1, 1)
texty    <- 9
textcex  <- 1.2
plotcex  <- 2.2
plotlwd  <- 3
titleline<- 0.4

## Panel 1
x.axis  <- c(-1, 0, 1)
textx   <- (max(x.axis)+min(x.axis))/2
coef.vec<- c(-0.176, -0.027, 0.227, 0.339, -0.136, -0.135, -0.342, 0.008)
se.vec  <- c(0.041, 0.003, 0.014, 0.165, 0.138, 0.095, 0.047, 0.002)
var.names <- c("Female","Age","Education","Jewish","Asian","Atheist","Muslim","Yrs France")
y.axis  <- length(coef.vec):1

par(mar=margins)
plot(coef.vec, y.axis, type="p", axes=FALSE, xlab="Regression Coefficient", ylab="",
     pch=19, cex=plotcex, xlim=c(min(x.axis), max(x.axis)), ylim=c(0,10), xaxs="r")
abline(h=y.axis, lty=2, lwd=.5, col="light grey")
segments(coef.vec-qnorm(.975)*se.vec, y.axis, coef.vec+qnorm(.975)*se.vec, y.axis, lwd=plotlwd)
segments(coef.vec-qnorm(.95)*se.vec,  y.axis-.1, coef.vec-qnorm(.95)*se.vec,  y.axis+.1, lwd=1.8, col="grey25")
segments(coef.vec+qnorm(.95)*se.vec,  y.axis-.1, coef.vec+qnorm(.95)*se.vec,  y.axis+.1, lwd=1.8, col="grey25")
axis(1, at=x.axis, labels=x.axis, cex.axis=1.2, mgp=c(2,.7,0))
axis(2, at=y.axis, labels=var.names, las=1, cex.axis=1.2, font=3, mgp=c(2,.6,0))
segments(0,0,0,8.5,lty=2)
segments(min(x.axis), texty-titleline, max(x.axis), texty-titleline, lty=1)
text(textx, texty, "Model (1) \nWithout Country F.E.", font=2, cex=textcex)

## Panel 2
x.axis  <- c(-1, 0, 1)
textx   <- (max(x.axis)+min(x.axis))/2
coef.vec<- c(-0.186, -0.030, 0.230, 0.499, -0.396, -0.082, -0.173, 0.008)
se.vec  <- c(0.042, 0.003, 0.017, 0.202, 0.208, 0.106, 0.107, 0.002)
y.axis  <- length(coef.vec):1

par(mar=margins2)
plot(coef.vec, y.axis, type="p", axes=FALSE, xlab="Regression Coefficient", ylab="",
     pch=19, cex=plotcex, xlim=c(min(x.axis), max(x.axis)), ylim=c(0,10), xaxs="r")
abline(h=y.axis, lty=2, lwd=.5, col="light grey")
segments(coef.vec-qnorm(.975)*se.vec, y.axis, coef.vec+qnorm(.975)*se.vec, y.axis, lwd=plotlwd)
segments(coef.vec-qnorm(.95)*se.vec,  y.axis-.1, coef.vec-qnorm(.95)*se.vec,  y.axis+.1, lwd=1.8, col="grey25")
segments(coef.vec+qnorm(.95)*se.vec,  y.axis-.1, coef.vec+qnorm(.95)*se.vec,  y.axis+.1, lwd=1.8, col="grey25")
axis(1, at=x.axis, labels=x.axis, cex.axis=1.2, mgp=c(2,.7,0))
segments(0,0,0,8.5,lty=2)
segments(min(x.axis), texty-titleline, max(x.axis), texty-titleline, lty=1)
text(textx, texty, "Model (2) \nWith Country F.E.", font=2, cex=textcex)

dev.off()

```

```{r}

## FIGURE 2a  (barplot response rates)
jpeg(
  filename = file.path(dir_figures, "ALVPNASCVResultsFig2a.jpg"),
  width = 2000, height = 2000, quality = 100, res = 300
)

height <- matrix(c(0.21, 0.27, 0.08, 0.25), nrow=2, ncol=2)

barplot(height, beside=TRUE, col=c("gray55","gray25"),
        ylim=c(0,0.4),
        names.arg=c("Marie Diouf vs. Aurélie Ménard",
                    "Khadija Diouf vs. Aurélie Ménard"),
        ylab="Positive response rate",
        main="Results from the C.V experiment on job discrimination in France\n based on religious cues",
        cex.lab=1.2, cex.axis=1.2)

legend(3.5, 0.35,
       c("Marie/Khadija Diouf","Aurélie Ménard"),
       fill=c("gray55","gray25"),
       cex=1.3)

dev.off()


```


```{r}

## FIGURE 2  (CI plot of differences)
jpeg(
  filename = file.path(dir_figures, "ALVPNASCVResultsGraph2.jpg"),
  width = 3000, height = 2000, quality = 100, res = 300
)

coef.vec <- c(5.79751, 16.54135)
se.vec   <- c(5.14188, 4.45876)

margins   <- c(2, 14, 0, 0)  # bottom, left, top, right
texty     <- 2.3
textcex   <- 1.1
plotcex   <- 1.5
plotlwd   <- 1
titleline <- 0.05

x.axis <- c(-10, 0, 30)
textx  <- (max(x.axis)+min(x.axis))/2

var.names <- c("... Aurélie Ménard vs. Marie Diouf",
               "...Aurélie Ménard vs. Khadija Diouf")
y.axis <- length(coef.vec):1

par(mar=margins)

plot(coef.vec, y.axis, type="p", axes=FALSE, xlab="", ylab="",
     pch=19, cex=plotcex, xlim=c(min(x.axis), max(x.axis)),
     ylim=c(0.5, length(coef.vec)+0.8), xaxs="r", col="grey25")

abline(h=y.axis, lty=2, lwd=.5, col="light grey")
segments(coef.vec-qnorm(.975)*se.vec, y.axis, coef.vec+qnorm(.975)*se.vec, y.axis, col="grey25", lwd=plotlwd)
segments(coef.vec-qnorm(.95)*se.vec,  y.axis-.05, coef.vec-qnorm(.95)*se.vec,  y.axis+.05, lwd=1.3, col="grey25")
segments(coef.vec+qnorm(.95)*se.vec,  y.axis-.05, coef.vec+qnorm(.95)*se.vec,  y.axis+.05, lwd=1.3, col="grey25")

axis(1, at=x.axis, labels=x.axis, cex.axis=1, mgp=c(2,.7,0))
axis(2, at=y.axis, labels=var.names, las=1, cex.axis=1, font=3, mgp=c(2,.6,0))
segments(0,0,0,length(coef.vec)+0.15,lty=2)

segments(min(x.axis), texty-titleline, max(x.axis), texty-titleline, lty=1)
text(textx, texty, "Difference in positive response rate for...", font=4, cex=textcex)

dev.off()


```

```{r}

## FIGURE 3  (CI plot for income determinants)
jpeg(
  filename = file.path(dir_figures, "PNAS2010ALVTable3.jpg"),
  width = 3700, height = 2000, quality = 100, res = 300
)

margins   <- c(4, 11, 0, 0)
texty     <- 5
textcex   <- 1.2
plotcex   <- 2.2
plotlwd   <- 3
titleline <- 0.4

x.axis   <- c(-0.1, 0, 0.7)
textx    <- (max(x.axis)+min(x.axis))/2
coef.vec <- c(0.357, 0.293, 0.055, 0.022)
se.vec   <- c(0.127, 0.120, 0.025, 0.035)

var.names <- c("Christian hh", "Head of hh is male", "Head of hh's education", "Education first migrant")
y.axis    <- length(coef.vec):1

par(mar=margins)
plot(coef.vec, y.axis, type="p", axes=FALSE, xlab="Regression Coefficient", ylab="",
     pch=19, cex=plotcex, xlim=c(min(x.axis), max(x.axis)), ylim=c(0,5.5), xaxs="r")
abline(h=y.axis, lty=2, lwd=.5, col="light grey")
segments(coef.vec-qnorm(.975)*se.vec, y.axis, coef.vec+qnorm(.975)*se.vec, y.axis, lwd=plotlwd)
segments(coef.vec-qnorm(.95)*se.vec,  y.axis-.1, coef.vec-qnorm(.95)*se.vec,  y.axis+.1, lwd=1.8, col="grey25")
segments(coef.vec+qnorm(.95)*se.vec,  y.axis-.1, coef.vec+qnorm(.95)*se.vec,  y.axis+.1, lwd=1.8, col="grey25")

axis(1, at=x.axis, labels=x.axis, cex.axis=1.2, mgp=c(2,.7,0))
axis(2, at=y.axis, labels=var.names, las=1, cex.axis=1.2, font=3, mgp=c(2,.6,0))
segments(0,0,0,4.2,lty=2)

segments(min(x.axis), texty-titleline, max(x.axis), texty-titleline, lty=1)
text(textx, texty, "Determinants of household monthly income", font=2, cex=textcex)

dev.off()

```

```{r}

message("\n══════════════════════════════════════════")
message("Replication complete. Outputs saved to:")
message(" • CV tables     → ", file.path(dir_tables, "cv"))
message(" • Survey tables → ", file.path(dir_tables, "sv"))
message(" • Figures       → ", dir_figures)
message("══════════════════════════════════════════")

```
