---
title: "01_cv_analysis.Rmd"
output: html_document
date: "2026-02-23"
---
```{r setup, include=FALSE}
source(here::here("code", "00_setup.R"))
```


##PART 1 — CV (RÉSUMÉ AUDIT) EXPERIMENT
#  Original Stata file: 2009Dec17CLA_CV.do
#   Design: matched pairs of CVs sent to employers.
#    • "Test" candidate  = candidate 1 (Khadija or Marie)
#    • "Control" candidate = candidate 2 (Aurélie)
#    • MuslimPair = 1 when pair is Khadija vs Aurélie
#    • ChristianPair = 1 when pair is Marie vs Aurélie

#  Three codings of the outcome differ in what counts as "missing":
# Coding 1 (ISM-Corum): drop NRNR, NRNo, NoNR, NoNo, ""
# Coding 2: drop NRNR, NRNo, NoNR, "" (keep NoNo as 0)
# Coding 3: drop only "" (invalid email tests)

# ── 1a. Variable construction ────────────────────────────────────────────────
# Stata generates binary indicators from string variables for pair type, position, contract, and firm size.  
# To replicate this in R: case_when

```{r}


cv <- original_data %>%
  mutate(
    # --- normalize firm size ---
    firm_size = firm_size %>%
      as.character() %>%
      str_replace_all("\u00A0", " ") %>%
      str_squish() %>%
      stri_replace_all_regex("\\p{Cc}", "") %>%   
      stri_trans_general("Latin-ASCII") %>%
      na_if(""),

    # --- normalize position ---
    position_clean = position %>%
      as.character() %>%
      str_replace_all("\u00A0", " ") %>%
      str_squish() %>%
      stri_replace_all_regex("\\p{Cc}", "") %>%   
      stri_trans_general("Latin-ASCII") %>%
      na_if(""),

    # --- pull the position code (01/02/03/04) ---
    position_code = str_extract(position_clean, "^\\d{2}"),

    # Pair type
    MuslimPair    = as.integer(str_detect(pair, "Khadija")),
    ChristianPair = as.integer(str_detect(pair, "Marie")),

    # Job position dummies 
    Secretary           = as.integer(position_code == "01"),
    AssistantDirector   = as.integer(position_code == "02"),
    AssistantAccountant = as.integer(position_code == "03"),
    Accountant          = as.integer(position_code == "04"),

    # Contract type
    CDD = as.integer(contract_type == "CDD"),
    CDI = as.integer(contract_type == "CDI"),

    # Firm size dummies
    SmallFirm  = case_when(
      is.na(firm_size) ~ NA_integer_,
      str_starts(firm_size, "01_") ~ 1L,
      TRUE ~ 0L
    ),
    MediumFirm = case_when(
      is.na(firm_size) ~ NA_integer_,
      str_starts(firm_size, "02_") ~ 1L,
      TRUE ~ 0L
    ),
    LargeFirm  = case_when(
      is.na(firm_size) ~ NA_integer_,
      str_starts(firm_size, "03_") ~ 1L,
      TRUE ~ 0L
    )
  )

```


# ── 1b. Merge corrected result variable ──────────────────────────────────────
# Stata: joinby testnumber using NewAdidaCoding.dta
# The corrected "newresult" variable fixes errors in the original "result".

```{r}

cv <- cv %>%
  mutate(testnumber = str_trim(as.character(testnumber)))

newcoding <- newcoding %>%
  mutate(testnumber = str_trim(as.character(testnumber)))

cvj <- left_join(cv, newcoding, by = "testnumber") %>%
  mutate(newresult = str_trim(as.character(newresult)))


```

#── 1c. Helpers: pseudo-R2, robust logit, & tidy output ──────────────────────────────────
# Stata: logit Y X, robust  →  uses Huber-White (HC0) standard errors.

```{r}

tidy_rlogit <- function(formula, data, model_name) {
  mf <- model.frame(formula, data = data, na.action = na.omit) # exact estimation sample
  fit <- glm(formula, data = mf, family = binomial(link = "logit"))
  ct  <- lmtest::coeftest(fit, vcov. = sandwich::vcovHC(fit, type = "HC0"))

  # Fit null model explicitly on the same sampled data
  null_fit <- glm(update(formula, . ~ 1), data = mf, family = binomial(link = "logit"))

  ll_full <- as.numeric(logLik(fit))
  ll_null <- as.numeric(logLik(null_fit))
  pseudo_r2 <- 1 - (ll_full / ll_null)

  tibble::tibble(
    model     = model_name,
    nobs      = nobs(fit),
    pseudo_r2 = pseudo_r2,
    term      = rownames(ct),
    estimate  = ct[, "Estimate"],
    std.error = ct[, "Std. Error"],
    statistic = ct[, "z value"],
    p.value   = ct[, "Pr(>|z|)"]
  )
}

```

# ── 1d. Helper:  ───────────────────────────
# Returns the data with ControlWins, TestWins, and the interact term.

```{r}

build_outcomes <- function(data, missing_values) {
  data %>%
    mutate(
      ControlWins = case_when(
        newresult %in% c("NRYes", "NoYes", "YesYes") ~ 1,
        newresult %in% missing_values                 ~ NA_real_,
        TRUE                                          ~ 0),
      TestWins = case_when(
        newresult %in% c("YesNR", "YesNo", "YesYes") ~ 1,
        newresult %in% missing_values                 ~ NA_real_,
        TRUE                                          ~ 0),
      interact = MuslimPair * test_photo
    )
}

```

# ── 1e. Define the three codings ────────────────────────────────────────────

```{r}

coding_defs <- list(
  C1 = c("", "NRNR", "NRNo", "NoNR", "NoNo"),   
  C2 = c("", "NRNR", "NRNo", "NoNR"),            
  C3 = c("")                                      
)

```

# ── 1f. Run all analyses across the three codings ───────────────────────────
# For each coding we produce:
#   Summary statistics (overall, by MuslimPair, by ChristianPair)
#   Welch t-tests (TestWins by MuslimPair; ControlWins by MuslimPair)
#   10 logit models with robust SEs 

```{r}

for (cname in names(coding_defs)) {

  d      <- build_outcomes(cvj, coding_defs[[cname]])
  
  # Drop rows with NA in test_photo before fitting any models!
  d <- d %>% filter(!is.na(test_photo))
  
  suffix <- cname   

  d      <- build_outcomes(cvj, coding_defs[[cname]])
  suffix <- cname   

  # Summary statistics
stats <- list(
  n    = ~ sum(!is.na(.)),
  mean = ~ mean(., na.rm = TRUE),
  sd   = ~ sd(., na.rm = TRUE)
)

overall <- d %>%
  summarise(
    across(c(TestWins, ControlWins), stats, .names = "{.col}_{.fn}")) %>%
  mutate(pair = "All", .before = 1)

by_pair <- bind_rows(
  d %>% filter(MuslimPair    == 1) %>%
    summarise(pair = "Muslim",    across(c(TestWins, ControlWins), stats, .names = "{.col}_{.fn}")),
  d %>% filter(ChristianPair == 1) %>%
    summarise(pair = "Christian", across(c(TestWins, ControlWins), stats, .names = "{.col}_{.fn}"))
)

cv_out[[paste0("summary_", suffix)]] <- bind_rows(overall, by_pair)

  # ---- t-tests ----
  # Stata: ttest TestWins, by(MuslimPair) unpaired unequal
  # Stata: ttest ControlWins, by(MuslimPair) unpaired unequal
  # Stata: ttest TestWins==ControlWins if ChristianPair==1, unpaired unequal
  # Stata: ttest TestWins==ControlWins if MuslimPair==1, unpaired unequal

  d_chr <- d %>% filter(ChristianPair == 1, !is.na(TestWins), !is.na(ControlWins))
  d_mus <- d %>% filter(MuslimPair    == 1, !is.na(TestWins), !is.na(ControlWins))

  tt <- bind_rows(
    tidy(t.test(TestWins    ~ MuslimPair, data = d, var.equal = FALSE)) %>%
      mutate(test = "TestWins by MuslimPair"),
    tidy(t.test(ControlWins ~ MuslimPair, data = d, var.equal = FALSE)) %>%
      mutate(test = "ControlWins by MuslimPair"),
    tidy(t.test(d_chr$TestWins, d_chr$ControlWins, paired = FALSE, var.equal = FALSE)) %>%
      mutate(test = "Test==Control | Christian"),
    tidy(t.test(d_mus$TestWins, d_mus$ControlWins, paired = FALSE, var.equal = FALSE)) %>%
      mutate(test = "Test==Control | Muslim")
  )

  cv_out[[paste0("ttests_", suffix)]] <- tt

# ---- Logit models ----

tidy_rlogit_m9_stata <- function(data, suffix) {

  data <- data %>% mutate(interact = MuslimPair * test_photo)

  f_full <- ControlWins ~ MuslimPair + CDD + SmallFirm + MediumFirm +
    Secretary + AssistantDirector + AssistantAccountant + test_photo + interact

  d_int <- data %>% filter(!is.na(interact), interact != 0, !is.na(ControlWins))
  sep_hit <- nrow(d_int) > 0 && (all(d_int$ControlWins == 1) || all(d_int$ControlWins == 0))

  # helper: McFadden pseudo-R2 on same estimation sample as fit
  # null_fit should use exact same model.frame as fit
  mcfadden_same_sample <- function(fit) {
    # Extract the exact estimation sample used by fit (accounts for all NA dropping)
    mf <- model.frame(fit)
    # Fit null model on the same sample, not on original data
    null_fit <- glm(update(formula(fit), . ~ 1), 
                    data = mf, 
                    family = binomial(link = "logit"))
    
    ll_full  <- as.numeric(stats::logLik(fit))
    ll_null  <- as.numeric(stats::logLik(null_fit))
    1 - (ll_full / ll_null)
  }

  if (sep_hit) {
    # drop interact term & drop interact!=0 observations
    d_fit <- data %>% filter(is.na(interact) | interact == 0)

    f_drop <- ControlWins ~ MuslimPair + CDD + SmallFirm + MediumFirm +
      Secretary + AssistantDirector + AssistantAccountant + test_photo

    # Create model.frame first with all variables in formula
    # Just to ensure both full and null models see identical sample
    mf_fit <- model.frame(f_drop, data = d_fit, na.action = na.omit)
    
    fit <- suppressWarnings(glm(f_drop, 
                                data = mf_fit, 
                                family = binomial(link = "logit")))
    ct  <- lmtest::coeftest(fit, vcov. = sandwich::vcovHC(fit, type = "HC0"))

    # McFadden on same sample
    pseudo_r2 <- mcfadden_same_sample(fit)

    res <- tibble::tibble(
      model     = paste0(suffix, "_m9_CW_interact"),
      nobs      = stats::nobs(fit),
      pseudo_r2 = pseudo_r2,
      term      = rownames(ct),
      estimate  = ct[, "Estimate"],
      std.error = ct[, "Std. Error"],
      statistic = ct[, "z value"],
      p.value   = ct[, "Pr(>|z|)"],
      note      = NA_character_
    ) %>%
      dplyr::bind_rows(
        tibble::tibble(
          model     = paste0(suffix, "_m9_CW_interact"),
          nobs      = stats::nobs(fit),
          pseudo_r2 = pseudo_r2,
          term      = "interact",
          estimate  = NA_real_, std.error = NA_real_,
          statistic = NA_real_, p.value = NA_real_,
          note      = "dropped: interact!=0 obs removed"
        )
      )

    return(res)
  }

  # full model branch
  
  mf_fit <- model.frame(f_full, data = data, na.action = na.omit)
  
  fit <- suppressWarnings(glm(f_full, 
                              data = mf_fit, 
                              family = binomial(link = "logit")))
  ct  <- lmtest::coeftest(fit, vcov. = sandwich::vcovHC(fit, type = "HC0"))

  pseudo_r2 <- mcfadden_same_sample(fit)

  tibble::tibble(
    model     = paste0(suffix, "_m9_CW_interact"),
    nobs      = stats::nobs(fit),
    pseudo_r2 = pseudo_r2,
    term      = rownames(ct),
    estimate  = ct[, "Estimate"],
    std.error = ct[, "Std. Error"],
    statistic = ct[, "z value"],
    p.value   = ct[, "Pr(>|z|)"],
    note      = NA_character_
  )
}

tidy_rlogit <- function(formula, data, model_name) {
  mf <- model.frame(formula, data = data, na.action = na.omit)
  # Fit full model on this controlled sample
  fit <- glm(formula, data = mf, family = binomial(link = "logit"))
  ct  <- lmtest::coeftest(fit, vcov. = sandwich::vcovHC(fit, type = "HC0"))
  # Fit null model on the same sample (not on original data!)
  null_fit <- glm(update(formula, . ~ 1), 
                  data = mf, 
                  family = binomial(link = "logit"))

  # McFadden pseudo-R2: both LLs computed on same N
  ll_full <- as.numeric(stats::logLik(fit))
  ll_null <- as.numeric(stats::logLik(null_fit))
  pseudo_r2 <- 1 - (ll_full / ll_null)

  tibble::tibble(
    model     = model_name,
    nobs      = stats::nobs(fit),
    pseudo_r2 = pseudo_r2,
    term      = rownames(ct),
    estimate  = ct[, "Estimate"],
    std.error = ct[, "Std. Error"],
    statistic = ct[, "z value"],
    p.value   = ct[, "Pr(>|z|)"]
  )
}

logit_results <- dplyr::bind_rows(

  tidy_rlogit(ControlWins ~ MuslimPair + CDD + SmallFirm + MediumFirm,
              d, paste0(suffix, "_m1_CW_base")),
  tidy_rlogit(ControlWins ~ MuslimPair + CDD + SmallFirm + MediumFirm + test_photo,
              d, paste0(suffix, "_m2_CW_photo")),

  tidy_rlogit(TestWins ~ MuslimPair + CDD + SmallFirm + MediumFirm,
              d, paste0(suffix, "_m3_TW_base")),
  tidy_rlogit(TestWins ~ MuslimPair + CDD + SmallFirm + MediumFirm + test_photo,
              d, paste0(suffix, "_m4_TW_photo")),

  tidy_rlogit(ControlWins ~ MuslimPair + CDD + SmallFirm + MediumFirm +
                Secretary + AssistantDirector + AssistantAccountant,
              d, paste0(suffix, "_m5_CW_pos")),
  tidy_rlogit(ControlWins ~ MuslimPair + CDD + SmallFirm + MediumFirm +
                Secretary + AssistantDirector + AssistantAccountant + test_photo,
              d, paste0(suffix, "_m6_CW_pos_photo")),

  tidy_rlogit(TestWins ~ MuslimPair + CDD + SmallFirm + MediumFirm +
                Secretary + AssistantDirector + AssistantAccountant,
              d, paste0(suffix, "_m7_TW_pos")),
  tidy_rlogit(TestWins ~ MuslimPair + CDD + SmallFirm + MediumFirm +
                Secretary + AssistantDirector + AssistantAccountant + test_photo,
              d, paste0(suffix, "_m8_TW_pos_photo")),

  tidy_rlogit_m9_stata(d, suffix),

  tidy_rlogit(TestWins ~ MuslimPair + CDD + SmallFirm + MediumFirm +
                Secretary + AssistantDirector + AssistantAccountant +
                test_photo + interact,
              d, paste0(suffix, "_m10_TW_interact"))
)

cv_out[[paste0("logits_", suffix)]] <- logit_results

cv_out[[paste0("logit_fit_", suffix)]] <- logit_results %>%
  dplyr::distinct(model, nobs, pseudo_r2) %>%
  dplyr::arrange(model)}

```

##Save results: 

```{r}

save_all_tables_from_list(
  out_list = cv_out,
  prefix   = "cv",
  base_dir = dir_tables,
  rows_per_page = 20,
  width = 2400, height = 1600, res = 200, digits = 4
)

```

