---
title: "02_survey_analysis"
output: html_document
date: "2026-02-23"
---

```{r setup, include=FALSE}
source(here::here("code", "00_setup.R"))
```

#PART 2 — SURVEY RESULTS 
#  Original Stata file: SurveyDataForPNAS.do
#  Three analyses:
#    A. OLS regression of income on first-migrant education, self-reported
#       religion, gender, age  (robust SEs)
#    B. "Table 2" ordered probit (david_religion, with interaction)
#    C. "Table A4" ordered probit + predicted probabilities for
#       typical Muslim vs typical Christian

# ── 2a. Variable construction ───────────────────────


```{r}

sv <- survey_data %>%
  mutate(
    # Gender: 0 = male, 1 = female (Stata recodes sexe 1→0, 2→1)
    gender = case_when(sexe == 1 ~ 0L, sexe == 2 ~ 1L, TRUE ~ NA_integer_),
    female = as.integer(gender == 1),
    Age    = as.numeric(age1),

    # Education of first migrant (99 = missing in Stata)
    educ_firstmigrant = if_else(q15 == 99, NA_real_, as.numeric(q15)),

    # Self-reported religion: 0 = Muslim, 1 = Christian
    self_religion     = case_when(q53 == 1 ~ 0, q53 == 2 ~ 1, TRUE ~ NA_real_),

    # Income (99 = missing)
    income     = if_else(q95 == 99, NA_real_, as.numeric(q95)),
    income_ord = ordered(if_else(q95 == 99, NA_integer_, as.integer(q95))),

    # "david_religion": both family & self must agree
    david_religion = case_when(
      q7 == 1 & q53 == 1 ~ 0,
      q7 == 2 & q53 == 2 ~ 1,
      TRUE ~ NA_real_),

    # Family religion
    christianfamily = case_when(q7 == 2 ~ 1, q7 == 1 ~ 0, TRUE ~ NA_real_),

    # Head-of-household gender
    genderhead = case_when(
      (q40 == "01" & female == 0) | q40 == "03" ~ 1,
      (q40 == "01" & female == 1) | q40 == "04" ~ 0,
      TRUE ~ NA_real_),

    # Education variables
    curreduc   = if_else(q43 < 99, as.numeric(q43), NA_real_),
    educfather = if_else(q36 == 99, NA_real_, as.numeric(q36)),
    educmother = if_else(q37 == 99, NA_real_, as.numeric(q37)),

    # Education of household head
    educhead = case_when(
      q40 == "01" ~ curreduc,
      q40 == "03" ~ educfather,
      q40 == "04" ~ educmother,
      TRUE ~ NA_real_),

    initialeduc = if_else(q15 < 99, as.numeric(q15), NA_real_)
  )

#helper functions: 

pseudo_r2_mcfadden_clm <- function(fit) {
  # use same estimation sample
  mf <- model.frame(fit)
  y  <- all.vars(formula(fit))[1]
  null_form <- stats::as.formula(paste0(y, " ~ 1"))
  null_fit  <- ordinal::clm(null_form, data = mf, link = "probit")

  1 - as.numeric(logLik(fit)) / as.numeric(logLik(null_fit))
}

# Robust SEs for ordinal::clm 
vcov_robust_clm_stata <- function(fit) {
  # Score contributions (n x k)
  S <- as.matrix(sandwich::estfun(fit))

  # Hessian of the log-likelihood (k x k)
  H <- fit$Hessian

  B <- solve(H)
  meat <- crossprod(S)
  V <- B %*% meat %*% B
  V
}

tidy_robust_clm_stata <- function(fit, model_name) {
  V  <- vcov_robust_clm_stata(fit)
  ct <- lmtest::coeftest(fit, vcov. = V)

  tibble::tibble(
    model     = model_name,
    term      = rownames(ct),
    estimate  = ct[, 1],
    std.error = ct[, 2],
    statistic = ct[, 3],
    p.value   = ct[, 4]
  )
}

```

# ── 2b. OUTPUT A: OLS regression ────────────────────────────────────────────
# Stata: reg income educ_firstmigrant self_religion gender Age, robust

```{r}

m_ols      <- lm(income ~ educ_firstmigrant + self_religion + gender + Age, data = sv)
ols_robust <- tidy(coeftest(m_ols, vcov. = vcovHC(m_ols, type = "HC0")))
sv_out[["OLS_robust"]] <- ols_robust

sv_out[["OLS_summary_stats"]] <- sv %>%
  summarise(across(c(income, educ_firstmigrant, self_religion, gender, Age),
                   list(n    = ~ sum(!is.na(.)),
                        mean = ~ mean(., na.rm = TRUE),
                        sd   = ~ sd(., na.rm = TRUE)),
                   .names = "{.col}_{.fn}"))

ols_fit_stats <- tibble(
  model      = "OLS_robust",
  n_obs      = nobs(m_ols),
  r2         = summary(m_ols)$r.squared,
  adj_r2     = summary(m_ols)$adj.r.squared
)

sv_out[["OLS_robust"]] <- tidy(coeftest(m_ols, vcov. = vcovHC(m_ols, type = "HC0"))) %>%
  mutate(n_obs = nobs(m_ols))

```

# ── 2c. OUTPUT B: Table 2 — Ordered probit (david_religion) ─────────────────
# Stata:
#   oprobit q95 q15 gender age1 david_religion if q95~=99 & q15~=99, robust
#   gen interaction = david_religion * q15
#   oprobit q95 q15 gender age1 david_religion interaction if …, robust

```{r}

sv_t2 <- sv %>%
  dplyr::filter(!is.na(income_ord), !is.na(educ_firstmigrant)) %>%
  dplyr::mutate(interaction = david_religion * educ_firstmigrant)

op1 <- ordinal::clm(
  income_ord ~ educ_firstmigrant + gender + Age + david_religion,
  data = sv_t2, link = "probit"
)

op2 <- ordinal::clm(
  income_ord ~ educ_firstmigrant + gender + Age + david_religion + interaction,
  data = sv_t2, link = "probit"
)

# robust coefficient tables
sv_out[["Table2_oprobit_robust"]] <- dplyr::bind_rows(
  tidy_robust_clm(op1, "Table2_col1"),
  tidy_robust_clm(op2, "Table2_col2_interaction")
)

# fit stats (N + McFadden pseudo-R2)
sv_out[["Table2_fit_stats"]] <- dplyr::bind_rows(
  tibble::tibble(
    model = "Table2_col1",
    n_obs = stats::nobs(op1),
    pseudo_r2_mcfadden = pseudo_r2_mcfadden_clm(op1)
  ),
  tibble::tibble(
    model = "Table2_col2_interaction",
    n_obs = stats::nobs(op2),
    pseudo_r2_mcfadden = pseudo_r2_mcfadden_clm(op2)
  )
)


```

# ── 2d. OUTPUT C: Table A4 — Ordered probit + predicted probabilities ────────
# Stata restricts to head-of-household respondents (q40 in {01, 03, 04}) and estimates: oprobit income christianfamily genderhead educhead initialeduc, robust
# Then computes predicted probabilities for:
#   (i)   typical Muslim vs. Christian by group-specific modes
#   (ii)  typical person (overall modes), varying only religion
#   (iii) typical person within ±0.5 SD of each covariate mean

```{r}

svA4 <- sv %>%
  filter(q40 %in% c("01", "03", "04")) %>%
  filter(!is.na(income), !is.na(christianfamily),
         !is.na(genderhead), !is.na(educhead), !is.na(initialeduc)) %>%
  mutate(income_ord = ordered(income))

opA4 <- polr(income_ord ~ christianfamily + genderhead + educhead + initialeduc,
             data = svA4, method = "probit", Hess = TRUE)

sv_out[["TableA4_oprobit"]] <- broom::tidy(opA4)
sv_out[["TableA4_fit_stats"]] <- tibble::tibble(
  model = "TableA4",
  n_obs = nobs(opA4),
  pseudo_r2_mcfadden = pseudo_r2_mcfadden(opA4)
)


# Predicted probabilities
pmat  <- predict(opA4, newdata = svA4, type = "probs")
pcols <- paste0("p", seq_len(ncol(pmat)))
svA4p <- bind_cols(svA4, as.data.frame(pmat) %>% setNames(pcols))

# Helper: statistical mode
Mode <- function(x) {
  x  <- x[!is.na(x)]
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# Helper: mean predicted probs for a subset
mean_probs <- function(df, label) {
  df %>%
    summarise(across(all_of(pcols), ~ mean(., na.rm = TRUE)), n = n()) %>%
    mutate(group = label, .before = 1)
}

# (i) Group-specific modes
modes_mus <- svA4p %>% filter(christianfamily == 0) %>%
  summarise(across(c(genderhead, educhead, initialeduc), Mode))
modes_chr <- svA4p %>% filter(christianfamily == 1) %>%
  summarise(across(c(genderhead, educhead, initialeduc), Mode))

pp_groupmode <- bind_rows(
  svA4p %>%
    filter(christianfamily == 0,
           genderhead  == modes_mus$genderhead,
           educhead    == modes_mus$educhead,
           initialeduc == modes_mus$initialeduc) %>%
    mean_probs("Muslim (group mode)"),
  svA4p %>%
    filter(christianfamily == 1,
           genderhead  == modes_chr$genderhead,
           educhead    == modes_chr$educhead,
           initialeduc == modes_chr$initialeduc) %>%
    mean_probs("Christian (group mode)")
)

# (ii) Overall modes, varying only religion
modes_all <- svA4p %>%
  summarise(across(c(genderhead, educhead, initialeduc), Mode))

pp_overallmode <- bind_rows(
  svA4p %>%
    filter(christianfamily == 0,
           genderhead  == modes_all$genderhead,
           educhead    == modes_all$educhead,
           initialeduc == modes_all$initialeduc) %>%
    mean_probs("Muslim (overall mode)"),
  svA4p %>%
    filter(christianfamily == 1,
           genderhead  == modes_all$genderhead,
           educhead    == modes_all$educhead,
           initialeduc == modes_all$initialeduc) %>%
    mean_probs("Christian (overall mode)")
)

# (iii) Mean ± 0.5 SD bands (overall)
band_filter <- function(df, ref_df = df) {
  stats <- ref_df %>%
    summarise(
      educhead_m    = mean(educhead, na.rm = TRUE),
      educhead_s    = sd(educhead, na.rm = TRUE),
      initialeduc_m = mean(initialeduc, na.rm = TRUE),
      initialeduc_s = sd(initialeduc, na.rm = TRUE)
    )

  df %>%
    filter(
      between(educhead,
              stats$educhead_m    - 0.5 * stats$educhead_s,
              stats$educhead_m    + 0.5 * stats$educhead_s),
      between(initialeduc,
              stats$initialeduc_m - 0.5 * stats$initialeduc_s,
              stats$initialeduc_m + 0.5 * stats$initialeduc_s)
    )
}

pp_band <- bind_rows(
  svA4p %>% filter(christianfamily == 0) %>% band_filter() %>% mean_probs("Muslim (±0.5 SD, group)"),
  svA4p %>% filter(christianfamily == 1) %>% band_filter() %>% mean_probs("Christian (±0.5 SD, group)")
)

pp_band_overall <- band_filter(svA4p, svA4p) %>%
  dplyr::group_by(christianfamily) %>%
  dplyr::summarise(dplyr::across(all_of(pcols), ~ mean(.x, na.rm = TRUE)),
                   n = dplyr::n(),
                   .groups = "drop") %>%
  dplyr::mutate(group = dplyr::if_else(christianfamily == 1,
                                      "Christian (±0.5 SD, overall)",
                                      "Muslim (±0.5 SD, overall)")) %>%
  dplyr::select(group, dplyr::everything(), -christianfamily)

sv_out[["TableA4_predprob_groupmode"]]    <- pp_groupmode
sv_out[["TableA4_predprob_overallmode"]]  <- pp_overallmode
sv_out[["TableA4_predprob_band_group"]]   <- pp_band
sv_out[["TableA4_predprob_band_overall"]] <- pp_band_overall

a4_fit_stats <- tibble(
  model = "TableA4",
  n_obs = nobs(opA4),
  logLik = as.numeric(logLik(opA4)),
  pseudo_r2_mcfadden = pseudo_r2_mcfadden(opA4)
)

svA4 <- sv %>%
  dplyr::filter(q40 %in% c("01", "03", "04")) %>%
  dplyr::filter(!is.na(income), !is.na(christianfamily),
                !is.na(genderhead), !is.na(educhead), !is.na(initialeduc)) %>%
  dplyr::mutate(income_ord = ordered(income))

opA4 <- ordinal::clm(
  income_ord ~ christianfamily + genderhead + educhead + initialeduc,
  data = svA4, link = "probit"
)

sv_out[["TableA4_oprobit_robust"]] <- tidy_robust_clm_stata(opA4, "TableA4")

```

#Save results: 

```{r}
save_all_tables_from_list(
  out_list = sv_out,
  prefix = "sv",
  base_dir = dir_tables,   
  rows_per_page = 20,
  width = 2400,
  height = 1600,
  res = 200,
  digits = 4
)

sv_out <- sv_out[sort(names(sv_out))]

```
